---
title: Independence and the Multiplication Rule
title_meta: Chapter 2
description: In this chapter, you will learn about independence, conditional probability, and the multiplication rule.
---

---type:MultipleChoiceExercise lang:r xp:50 skills:1
## Exercise 1. Independence
Two events $A$ and $B$ are independent if $\mbox{Pr}(A \mbox{ and } B) = \mbox{Pr}(A) P(B)$. Under which situation are the draws independent?

A. You don't replace the draw. B. You replace the draw. C. Neither D. Both

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 2. Sampling with replacement
Say you’ve drawn 5 balls from the a box that has 3 cyan balls, 5 magenta balls, and 7 yellow balls, with replacement, and all have been yellow. What is the probability that the next one is yellow?

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 3. Sampling with replacement - generalized
Write exercise that has coding to calculate this probability in general

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 4. Rolling a die
If you roll a 6-sided die six times, what is the probability of not seeing a 6?

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 5. Probability the Celtics win a game
Two teams, say the Celtics and the Cavs, are playing a seven game series. The Cavs are a better team and have a 60% chance of winning each game. What is the probability that the Celtics win at least one game?

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 6. Monte Carlo simulation for Celtics winning a game
Create a Monte Carlo simulation to confirm your answer to the previous problem. Use `B <- 10000` simulations.

Hint: Use the following code to generate the results of the first four games:
```{r}
celtic_wins <- sample(c(0,1), 4, replace = TRUE, prob = c(0.6, 0.4))
```
The Celtics must win one of these 4 games, or the series will be over!
*** =sample_code
```{r}
celtic_wins <- sample(c(0,1), 4, replace = TRUE, prob = c(0.6, 0.4))
```

*** =solution
```{r}
B <- 10000
results <- replicate(B, {
  celtic_wins <- sample(c(0,1), 4, replace = TRUE, prob = c(0.6, 0.4))
  if(any(celtic_wins==1)) return(TRUE) else return(FALSE)
})
mean(results)
```
