---
title: The Central Limit Theorem
title_meta: Chapter 6
description: In this chapter, you will learn about the Central Limit Theorem.
---

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 1. American Roulette probability of winning money
In the previous assessment, you created a random variable $S$ that is the sum of your winnings after betting on green 1000 times.

What is the probability that you end up winning money? Hint: use the CLT.

*** =solution
```{r}
avg <- n * (17*1/19 + -1*18/19)
se <- sqrt(n) * (17 - -1)*sqrt(1/19*18/19)

1 - pnorm(0, avg, se)
```

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 2. American Roulette Monte Carlo simulation
Create a Monte Carlo simulation that generates 10000 outcomes of $S$. Compute the average and standard deviation of the resulting list to confirm the results of 6 and 7. Start your code by setting the seed to 1 with `set.seed(1)`.

*** =solution
```{r}
set.seed(1)
B <- 10000
S <- replicate(B,{  
  X <- sample(c(17,-1), size = n, replace = TRUE, prob = c(1/19, 18/19))
  sum(X)
})
mean(S)
sd(S)
```

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 3. American Roulette Monte Carlo vs CLT
Now check your answer to 8 using the Monte Carlo result.
*** =solution
```{r}
mean(S>0)
```

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 4. American Roulette Monte Carlo vs CLT comparison
CONVERT TO MULTIPLE CHOICE
The Monte Carlo result and the CLT approximation are close, but not that close. What could account for this?
A. 10,000 simulations is not enough. If we do more, they match. B. The CLT does not work as well when the probability of success is small. In this case, it was 1/19. If we make the number of roulette plays bigger, they will match better. C. The difference is within rounding error. D. The CLT only works for the averages.

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 5. American Roulette average winnings per bet

Now create a random variable $Y$ that is your average winnings per bet after playing off your winnings after betting on green 10,000 times.

*** =solution
```{r}
    set.seed(1)
    n <- 10000
    X <- sample(c(17,-1), size = n, replace = TRUE, prob = c(1/19, 18/19))
    Y <- mean(X)
```

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 6. American Roulette per bet expected value
What is the expected value of $Y$?
*** =solution
17*1/19 + -1*18/19

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 7. American Roulette per bet standard error
What is the standard error of $Y$?
*** =solution
(17 - -1)*sqrt(1/19*18/19)/sqrt(n) 

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 8. American Roulette winnings per game are positive
What is the probability that you end up winnings per game are positive? Hint: use the CLT.

*** =solution
```{r}
avg <- 17*1/19 + -1*18/19
se <- 1/sqrt(n) * (17 - -1)*sqrt(1/19*18/19)

1 - pnorm(0, avg, se)
```

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 9. American Roulette Monte Carlo again
Create a Monte Carlo simulation that generates 2,500 outcomes of $S$. Compute the average and standard deviation of the resulting list to confirm the results of 6 and 7. Start your code by setting the seed to 1 with set.seed(1)

*** =solution
```{r}
set.seed(1)
n <- 10000
B <- 2500
S <- replicate(B,{  
  X <- sample(c(17,-1), size = n, replace = TRUE, prob = c(1/19, 18/19))
  mean(X)
})
mean(S)
sd(S)
```
---type:NormalExercise lang:r xp:100 skills:1
## Exercise 10. American Roulette comparison
Now check your answer to exercise 8 using the Monte Carlo result.
*** =solution
mean(S>0)

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 11. American Roulette comparison analysis
CHANGE TO MULTIPLE CHOICE
The Monte Carlo result and the CLT approximation are now much closer. What could account for this?
A. We are now computing averages instead of sums. B. 2,500 Monte Carlo simulations is not better than 10,000. D. The CLT does works better when the sample size is larger. We increased from 1,000 to 10,000. D. It is not closer. The difference is within rounding error.
