---
title: The Addition Rule
title_meta: Chapter 3
description: In this chapter, you will learn about the addition rule and the Monty Hall problem.
---

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 1. The Cavs and the Warriors
Two teams, say the Cavs and the Warriors, are playing a seven game championship series. The first to win four games therefore wins the series. The teams are equally good so they each have a 50-50 chance of winning each game. If the Cavs lose the first game, what is the probability that they win the series?

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 2. The Cavs and the Warriors - Monte Carlo
Confirm the results of the previous question with a Monte Carlo simulation.
*** =solution
```{r}
B <- 100000
results <- replicate(B, {
  cavs_wins <- sample(c(0,1), 6, replace = TRUE)
  sum(cavs_wins)>=4 
})
mean(results)
```

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 3. A and B play a series - part 1
Two teams, $A$ and $B$, are playing a seven series game series. Team $A$ is better than team $B$ and has a $p>0.5$ chance of winning each game. Give a value $p$, the probability of winning the series for the underdog team $B$ can be computed with the following function based on a Monte Carlo simulation:

```{r}
prob_win <- function(p){
  B <- 10000
  result <- replicate(B, {
    b_win <- sample(c(1,0), 7, replace = TRUE, prob = c(1-p, p))
    sum(b_win)>=4
    })
  mean(result)
}
```
***= instructions
- Use the function `sapply` to compute the probability, call it `Pr` of winning for `p <- seq(0.5, 0.95, 0.025)`.
- Then plot the result `plot(p, prob)`.

***= solution
p <- seq(0.5, 0.95, 0.025)
Pr <- sapply(p, prob_win)
plot(p, Pr)

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 4. A and B play a series - part 2
Repeat the previous exercise, but now keep the probability fixed at `p <- 0.75` and compute the probability for different series lengths. Best of 1 game, 3 games, 5 games, .... Specifically, `N <- seq(1, 25, 2)`. Hint: use this function now.
    ```{r}
    prob_win <- function(N, p=0.75){
      B <- 10000
      result <- replicate(B, {
        b_win <- sample(c(1,0), N, replace = TRUE, prob = c(1-p, p))
        sum(b_win)>=(N+1)/2
        })
      mean(result)
    }
    ```
    
***= solution
```{r, eval=FALSE}
N <- seq(1, 25, 2)
Pr<- sapply(N, prob_win)
plot(N, Pr)
