---
title: The Addition Rule
title_meta: Chapter 3
description: In this chapter, you will learn about the addition rule and the Monty Hall problem.
---

## Exercise 1. The Cavs and the Warriors

```yaml
type: NormalExercise
lang: r
xp: 100
skills: 1
```
Two teams, say the Cavs and the Warriors, are playing a seven game championship series. The first to win four games wins the series. The teams are equally good, so they each have a 50-50 chance of winning each game. 

If the Cavs lose the first game, what is the probability that they win the series?

`@instructions`
- Assign the number of remaining games to the variable `n`.
- Use the `list` function to create a list of game outcomes, where 0 indicates a loss for the Cavs and 1 indicates a win for the Cavs. Assign this value to the variable `n`.
- Use the `expand.grid` function to create a data frame containing all the possibilities for outcomes of the remaining games.
- To use the `rep` function within the `expand.grid` function to indicate the number of columns the results data frame should contain.
- Use the `rowSums` function to identify which combinations of game outcomes result in the Cavs winning the number of games necessary to win the series.
- Use the `mean` function to calculate the proportion of outcomes that result in the Cavs winning the series. 

`@hint`
- Any combination of 4 or more wins across the 6 remaining games allows the Cavs to win the series.
- The results data frame should contain one column for each remaining game.
- The results data frame should contain one row for each combination of wins (1) or losses (0) for the next 6 games.

`@pre_exercise_code`
```{r}
# no pec
```

`@sample_code`
```{r}
# Assign a variable 'n' as the number of remaining games.


# Assign a variable 'l' to a list of possible game outcomes, where 0 indicates a loss and 1 indicates a win for the Cavs. 


# Create a data frame named 'possibilities' that contains all possible outcomes for the remaining games.


# Create a vector named 'results' that indicates whether each row in the data frame 'possibilities' contains enough wins for the Cavs to win the series.


# Calculate the proportion of 'results' in which the Cavs win the series. Print the outcome to the console.

```

`@solution`
```{r}
# Assign a variable 'n' as the number of remaining games.
n <- 6

# Assign a variable 'l' to a list of possible game outcomes, where 0 indicates a loss and 1 indicates a win for the Cavs. 
l <- list(0:1)

# Create a data frame named 'possibilities' that contains all possible outcomes for the remaining games.
possibilities <- expand.grid(rep(l, n))

# Create a vector named 'results' that indicates whether each row in the data frame 'possibilities' contains enough wins for the Cavs to win the series.
results <- rowSums(possibilities)>=4

# Calculate the proportion of 'results' in which the Cavs win the series. Print the outcome to the console.
mean(results)
```

`@sct`
```{r}
test_error()
test_output_contains("n <- 6", incorrect_msg = "Make sure you assign the number of games remaining in the seven game series to the variable `n`.")
test_output_contains("l <- list(0:1)", incorrect_msg = "Make sure you use the `mean` function on the `results` object.")
test_output_contains("possibilities <- expand.grid(rep(l, n))", incorrect_msg = "Make sure you create a data frame called 'possibilities' that uses the expand.grid function to replicate the list of outcomes over the number of remaining games.")
test_output_contains("results <- rowSums(possibilities)>=4", incorrect_msg = "Make sure you create a vector called 'results' that finds rows in the data frame 'possibilities' that contain enough won games for the Cavs to win the series.")
test_output_contains("0.34375", incorrect_msg = "You are not providing a calculation that gives the correct answer. Make sure you are finding possible series outcomes that contain at least 4 games won for the Cavs.")
success_msg("Great work! Let's move on to the next problem.")
```

## Exercise 2. The Cavs and the Warriors - Monte Carlo

```yaml
type: NormalExercise
lang: r
xp: 100
skills: 1
```

Confirm the results of the previous question with a Monte Carlo simulation to estimate the probability of the Cavs winning the series after losing the first game.

`@instructions`
- Use the `replicate` function to replicate the sample code for `B <- 10000` simulations.
- Use the `sample` function to simulate a series of 6 games with random, independent outcomes of either a loss for the Cavs (0) or a win for the Cavs (1).
- Use the `sum` function to indicate which of the simulated series contained at least 4 wins for the Cavs.
- Use the `mean` function to find the proportion of simulations in which the Cavs win at least 4 of the remaining games.

`@hint`
- To use the `replicate` function, provide the number of replications and the code you want to be replicated. You can use {} to include multiple expressions that you'd like to be repeated.
```{r}
results <- replicate(number_of_times_to_replicate, {
  first_command_to_run
  second_command_to_run
})
```

- You can `sample` function to simulate a series of 6 games where the outcome is either "0" for a Cav loss or "1" for a Cav win. The probabilities of each outcome are equal and independent.
```{r}
cavs_wins <- sample(c(0,1), 6, replace = TRUE)
```

`@pre_exercise_code`
```{r}
# no pec
```

`@sample_code`
```{r}
# The variable `B` specifies the number of times we want the simulation to run. Let's run the Monte Carlo simulation 10,000 times.
B <- 10000

# Use the `set.seed` function to make sure your answer matches the expected result after random sampling.
set.seed(1)

# Create an object called `results` that replicates the sample code for `B` iterations and tallies the number of simulated series that contain at least four wins for the Cavs.




# Calculate the frequency out of `B` iterations that the Cavs won at least four games in the remainder of the series. Print your answer to the console. 

```

`@solution`
```{r}
# The variable `B` specifies the number of times we want the simulation to run. Let's run the Monte Carlo simulation 10,000 times.
B <- 10000

# Use the `set.seed` function to make sure your answer matches the expected result after random sampling.
set.seed(1)

# Create an object called `results` that replicates the sample code for `B` iterations and tallies the number of simulated series that contain at least four wins for the Cavs.
results <- replicate(B, {
  cavs_wins <- sample(c(0,1), 6, replace = TRUE)
  sum(cavs_wins)>=4 
})

# Calculate the frequency out of `B` iterations that the Cavs won at least four games in the remainder of the series. Print your answer to the console. 
mean(results)
```

`@sct`
```{r}
test_error()
test_output_contains("sample(c(0,1), 6, replace = TRUE)", incorrect_msg = "Make sure you use the `sample` function to simulate random outcomes for a series of 6 games. Use c(0,1) to indicate that the sample should include a 0 for a loss or a 1 for a win.")
test_output_contains("sum(cavs_wins)>=4", incorrect_msg = "Make sure you use the `sum` function to generate results that indicate simulated series where the Cavs won 4 or more games in the series.")
test_output_contains("0.3453", incorrect_msg = "You are not providing a calculation that gives the correct answer. Make sure you've executed the `set.seed(1)` command and have replicated the sample code over 10,000 iterations.")
success_msg("Great work!")
```

## Exercise 3. A and B play a series - part 1

```yaml
type: NormalExercise
lang: r
xp: 100
skills: 1
```
Two teams, $A$ and $B$, are playing a seven series game series. Team $A$ is better than team $B$ and has a $p>0.5$ chance of winning each game.

`@instructions`
- Use the function `sapply` to compute the probability, call it `Pr` of winning for `p <- seq(0.5, 0.95, 0.025)`.
- Then plot the result `plot(p, prob)`.

`@hint`
- To use the `replicate` function, provide the number of replications and the code you want to be replicated. You can use {} to include multiple expressions that you'd like to be repeated.

`@pre_exercise_code`
```{r}
# no pec
```

`@sample_code`
```{r}
# Let's assign the variable 'p' as the vector of probabilities that team $A$ will win.
p <- seq(0.5, 0.95, 0.025)

# Given a value 'p', the probability of winning the series for the underdog team $B$ can be computed with the following function based on a Monte Carlo simulation:
prob_win <- function(p){
  B <- 10000
  result <- replicate(B, {
    b_win <- sample(c(1,0), 7, replace = TRUE, prob = c(1-p, p))
    sum(b_win)>=4
    })
  mean(result)
}

# Apply the 'prob_win' function across the vector of probabilities that team $A$ will win to determine the probability that team $B$ will win.


# Plot the probability 'p' on the x-axis and 'Pr' on the x-axis.

```

`@solution`
```{r}
# Let's assign the variable 'p' as the vector of probabilities that team $A$ will win.
p <- seq(0.5, 0.95, 0.025)

# Given a value 'p', the probability of winning the series for the underdog team $B$ can be computed with the following function based on a Monte Carlo simulation:
prob_win <- function(p){
  B <- 10000
  result <- replicate(B, {
    b_win <- sample(c(1,0), 7, replace = TRUE, prob = c(1-p, p))
    sum(b_win)>=4
    })
  mean(result)
}

# Apply the 'prob_win' function across the vector of probabilities that team $A$ will win to determine the probability that team $B$ will win.
Pr <- sapply(p, prob_win)

# Plot the probability 'p' on the x-axis and 'Pr' on the y-axis.
plot(p, Pr)
```

`@sct`
```{r}
test_error()
test_output_contains("sapply(p, prob_win)", incorrect_msg = "Make sure you use the `sample` function to apply the function `prob_win` function across the vector of probabilities that team $A$ will win.")
test_output_contains("plot(p, Pr)", incorrect_msg = "Make sure you use the `plot` function to generate the plot where 'p' is on the x-axis and 'Pr' is on the y-axis.")
success_msg("Great work! Let's move on to the next example.")
```

## Exercise 4. A and B play a series - part 2

```yaml
type: NormalExercise
lang: r
xp: 100
skills: 1
```
Repeat the previous exercise, but now keep the probability that team $A$ wins fixed at `p <- 0.75` and compute the probability for different series lengths. For example, wins in best of 1 game, 3 games, 5 games, and so on through a series that lasts 25 games.

`@instructions`
- Use the `seq` function to generate a list of odd numbers ranging from 1 to 25. 
- Use the function `sapply` to compute the probability, call it `Pr`, of winning during series of different lengths.
- Then plot the result `plot(p, prob)`.

`@hint`
- Be sure to execute the code that generates the `prob_win` function.

`@pre_exercise_code`
```{r}
# no pec
```

`@sample_code`
```{r}
# Given a value 'p', the probability of winning the series for the underdog team $B$ can be computed with the following function based on a Monte Carlo simulation:
prob_win <- function(N, p=0.75){
      B <- 10000
      result <- replicate(B, {
        b_win <- sample(c(1,0), N, replace = TRUE, prob = c(1-p, p))
        sum(b_win)>=(N+1)/2
        })
      mean(result)
    }

# Assign the variable 'N' as the vector of series lengths. Use only odd numbers ranging from 1 to 25 games.


# Apply the 'prob_win' function across the vector of series lengths to determine the probability that team $B$ will win.


# Plot the number of games in the series 'N' on the x-axis and 'Pr' on the x-axis.

```

`@solution`
```{r}
# Given a value 'p', the probability of winning the series for the underdog team $B$ can be computed with the following function based on a Monte Carlo simulation:
prob_win <- function(N, p=0.75){
      B <- 10000
      result <- replicate(B, {
        b_win <- sample(c(1,0), N, replace = TRUE, prob = c(1-p, p))
        sum(b_win)>=(N+1)/2
        })
      mean(result)
    }

# Assign the variable 'N' as the vector of series lengths. Use only odd numbers ranging from 1 to 25 games.
N <- seq(1, 25, 2)

# Apply the 'prob_win' function across the vector of series lengths to determine the probability that team $B$ will win.
Pr<- sapply(N, prob_win)

# Plot the number of games in the series 'N' on the x-axis and 'Pr' on the x-axis.
plot(N, Pr)
```

`@sct`
```{r}
test_error()
test_output_contains("sapply(N, prob_win)", incorrect_msg = "Make sure you use the `sample` function to apply the function `prob_win` function across the vector of series lengths.")
test_output_contains("plot(N, Pr)", incorrect_msg = "Make sure you use the `plot` function to generate the plot where 'N' is on the x-axis and 'Pr' is on the y-axis.")
success_msg("Great work!")
```
