---
title: The Big Short
title_meta: Chapter 7
description: In this chapter, you will learn about interest rates and the big short.
---

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 1. Bank earnings

Create a random variable $S$ with the earnings of your bank if you give out 10,000 loans, the default rate is 0.3, and you lose $200,000 in each foreclosure. Hint: use code we showed in the videos, but change the parameters.

*** =solution
```{r}
n <- 10000
loss_per_foreclosure <- -200000
p <- 0.03 
defaults <- sample( c(0,1), n, prob=c(1-p, p), replace = TRUE)
S <- sum(defaults * loss_per_foreclosure)
```

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 2. Bank earnings Monte Carlo
Run a Monte Carlo simulation with 10,000 outcomes for $S$. Make a histogram of the results.

*** =solution
```{r}
B <- 10000
S <- replicate(B, {
    defaults <- sample( c(0,1), n, prob=c(1-p, p), replace = TRUE) 
  sum(defaults * loss_per_foreclosure)
})
hist(S)
```
---type:NormalExercise lang:r xp:100 skills:1
## Exercise 3. Bank earnings expected value
What is the expected value of $S$:

*** =solution
n*(p*loss_per_foreclosure + (1-p)*0)

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 4. Bank earnings standard error
What is the standard error of $S$?

*** =solution
sqrt(n)*abs(loss_per_foreclosure)*sqrt(p*(1-p))

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 5. Bank earnings interest rate - 1
Suppose we give out loans for $180,000. What should the interest rate be so that our expected value is 0?

*** =solution
$$ lp + x(1-p) = 0 $$

This implies that $x$ is:

- loss_per_foreclosure*p/(1-p)
and our interest rate:

- p*loss_per_foreclosure/(1-p) / 180000

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 6. Bank earnings interest rate - 2
What should the interest rate be so that the chance of losing money is 1 in 20? In math notation, what should the interest rate be so that $\mbox{Pr}(S<0) = 0.05$?

*** =solution
Solution $$ \mbox{Pr}\left(\frac{S - \mbox{E}[S]}{\mbox{SE}[S]} < \frac{ - \mbox{E}[S]}{\mbox{SE}[S]}\right) $$

$$ \mbox{Pr}\left(Z < \frac{- { lp + x(1-p)}n}{(x-l) \sqrt{np(1-p)}}\right) = 0.01 $$

Let $z=$qnorm(0.05) gives us the value of $z$ for which:

$$ \mbox{Pr}(Z \leq z) = 0.05 $$

This means that right side of the complicated equation must be $z$=qnorm(0.01):

$$ \frac{- { lp + x(1-p)}N}{(x-l) \sqrt{Np(1-p)}} = z $$

We solve for $x$:

$$ x = - l \frac{ np - z \sqrt{np(1-p)}}{N(1-p) + z \sqrt{np(1-p)}}$$

which is:
```{r}
l <- loss_per_foreclosure
z <- qnorm(0.05)
x <- -l*( n*p - z*sqrt(n*p*(1-p)))/ ( n*(1-p) + z*sqrt(n*p*(1-p)))
x/180000
```

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 7. Bank earnings - minimize money loss
CHANGE TO MULTIPLE CHOICE
If the bank wants to minimize the probabilities of losing money, which of the following does not make interest rates go up?
A. A smaller pool of loans. B. A larger probability of default. C. A smaller required probability of losing money. D. The number of Monte Carlo simulations.
