---
title: Random Variables and Sampling Models
title_meta: Chapter 5
description: In this chapter, you will learn about random variables and sampling models
---

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 1. American Roulette probabilities

In American Roulette you can also bet on green. Remember there are 18 reds, 18 blacks and 2 greens (0 and 00). What are the chances the green comes out?
*** =solution
2/38 = 1/19

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 2. American Roulette payout
The payout for winning on green is $17 dollars. This means that if you bet a dollar and it lands on green you get $17. Create a sampling model using sample to simulate the random variable $X$ for your winnings. Hint: see the exampe below for how it should look like when betting on red.

```{r}
X <- sample(c(1,-1), 1, prob = c(9/19, 10/19))
```

*** =solution
X <- sample(c(17,-1), 1, prob = c(1/19, 18/19))

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 3. American Roulette expected value
Compute the expected value of $X$.

*** =solution
17*1/19 + -1*18/19

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 4. American Roulette standard error
Compute the standard error of $X$.

*** =solution
(17 - -1)*sqrt(1/19*18/19)

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 5. American Roulette sum of winnings
Now create a random variable $S$ that is the sum of your winnings after betting on green 1000 times. Hint: change the argument size and replace in your answer to exercise 2. Start your code by setting the seed to 1 with set.seed(1).

*** =solution
```{r}
    set.seed(1)
    n <- 1000
    X <- sample(c(17,-1), size = n, replace = TRUE, prob = c(1/19, 18/19))
    S <- sum(X)
 ```

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 6. American Roulette winnings expected value
What is the expected value of $S$?

*** =solution
n * (17*1/19 + -1*18/19)

---type:NormalExercise lang:r xp:100 skills:1
## Exercise 7. American Roulette winnings standard error
What is the standard error of $S$?

*** =solution
sqrt(n) * (17 - -1)*sqrt(1/19*18/19)
